---
title: "Tutorial on rmr2"
author: "Andrey Ziyatdinov"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: united
    toc: true
    keep_md: true
---

```{r options, echo = F}
opts_chunk$set(fig.path = "figures/", comment = NA, results = 'markup', tidy = F, message = F, warning = F, echo = T)
```

```{r inc, echo = F, cache = FALSE}
library(plyr)
library(ggplot2)
library(gridExtra)

library(pander)
```

```{r settings, echo = F, cache = FALSE}
theme_set(theme_light())
panderOptions('table.split.table', Inf)
panderOptions('knitr.auto.asis', FALSE)
```

# About

References

* Example 1 Squares of integers & Example 2 Groups: [basic-examples.R](https://github.com/RevolutionAnalytics/rmr2/blob/master/pkg/tests/basic-examples.R)
* Example 3 Words Count: [wordcount.R](https://github.com/RevolutionAnalytics/rmr2/blob/master/pkg/tests/wordcount.R)

# Preparation

Include the R-Hadoop libraries:

```{r inc_rhadoop, cache = FALSE}
library(rhdfs)
library(rhbase)
library(rmr2)
```

Set up Hadoop environment variables:

```{r env_hadoop}
Sys.setenv("HADOOP_CMD" = "/home/hduser/hadoop/bin/hadoop")
Sys.setenv("HADOOP_STREAMING" = "/home/hduser/hadoop/share/hadoop/tools/lib/hadoop-streaming-2.7.2.jar")
```

# Map-reduce Examples

## Map-reduce Example 1: Squares of integers

```{r ex1, cache = TRUE}
small.ints <- to.dfs(1:1000)

ret1 <- mapreduce(
  input = small.ints, 
  map = function(k, v) cbind(v, v^2))

out1 <- from.dfs(ret1)
```

Plot results:

```{r plot1, cache = TRUE}
df <- as.data.frame(out1$val)
names(df) <- c("v", "v2")

ggplot(df, aes(v, v2)) + geom_point()
```

## Map-reduce Example 2: Groups

Data in R:

```{r gr, cache = T}
set.seed(1)
groups <- rbinom(32, n = 50, prob = 0.4)
```

R code with `tapply` equivalent to map-reduce operation:

```{r tapply, cache = T}
tapply(groups, groups, length)
```

Send data to HDFS:

### Reduce is skipped

```{r groups_dfs, cache = TRUE}
groups_dfs <- to.dfs(groups)
```

```{r ex2_1, cache = TRUE}
ret2_1 <- mapreduce(
  input = groups_dfs, 
  map = function(k, v) keyval(v, 1))

out2_1 <- from.dfs(ret2_1)
```

Table of results (first 5 rows):

```{r tab2_1, echo = FALSE, results = 'asis'}
pander(head(as.data.frame(out2_1), 5), style = 'rmarkdown')
```

### Reduce is on

```{r ex2, cache = TRUE}
ret2_2 <- mapreduce(
  input = groups_dfs, 
  map = function(k, v) keyval(v, 1), 
  reduce = function(k, vv) keyval(k, length(vv)))

out2_2 <- from.dfs(ret2_2)
```

Table of results:

```{r tab2_2, echo = FALSE, results = 'asis'}
pander(as.data.frame(out2_2), style = 'rmarkdown')
```

## Map-reduce Example 3: Word counter

Word-count functions:

```{r fun3}
wc_map <- function(k, lines, pattern = " ") 
{
  keyval(
    unlist(strsplit(lines, pattern)), 
    1)
}

wc_reduce <- function(word, counts) 
{
  keyval(
    word, 
    sum(counts))
}
```

```{r text3, cache = T}
text <- capture.output(license())
text_dfs <- to.dfs(text)
```

```{r show_text}
text
```

```{r ex3, cache = T}
ret3 <- mapreduce(
  input = text_dfs, 
  map = wc_map, reduce = wc_reduce, combine = TRUE)
  
out3 <- from.dfs(ret3)  
```

Top 5 words:

```{r tab5, echo = FALSE, results = 'asis'}
df <- as.data.frame(out3)

ord <- order(df$val, decreasing = TRUE)
df <- df[ord, ]

rownames(df) <- NULL

pander(head(df, 5), style = 'rmarkdown')
```
