---
title: "RcppParallel Examples"
author: "Andrey Ziyatdinov"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: united
    toc: true
    keep_md: true
---

```{r options, echo = F}
opts_chunk$set(fig.path = "figures/", comment = NA, results = 'markup', tidy = F, message = F, warning = F, echo = T, cache = T)
```

```{r inc, echo = F, cache = FALSE}
library(plyr)
library(ggplot2)
library(gridExtra)

library(pander)
```

```{r settings, echo = F, cache = FALSE}
theme_set(theme_light())
panderOptions('table.split.table', Inf)
panderOptions('knitr.auto.asis', FALSE)
```

# About


## References

* Plot dendograms: [ggdendro vignette](https://cran.r-project.org/web/packages/ggdendro/vignettes/ggdendro.html)

# Include

```{r include, cache = F}
library(Rcpp)
library(RcppParallel)
library(RcppArmadillo)

library(microbenchmark)

library(ggdendro)
```

# Inner Product

The code is copied from [gallery.rcpp.org/articles/parallel-inner-product/](http://gallery.rcpp.org/articles/parallel-inner-product/).

## Serial version

```{r ip_serial}
sourceCpp(code = '
  #include <Rcpp.h>
  using namespace Rcpp;

  #include <algorithm>

  // [[Rcpp::export]]
  double innerProduct(NumericVector x, NumericVector y) {
     return std::inner_product(x.begin(), x.end(), y.begin(), 0.0);
  }'
)
```  

## Parallel version

```{r ip_par1}
sourceCpp(code = '
  #include <Rcpp.h>
  #include <RcppParallel.h>

  // [[Rcpp::depends(RcppParallel)]]
  using namespace RcppParallel;
  using namespace Rcpp;
  
  using namespace RcppParallel;

  struct InnerProduct : public Worker
  {   
     // source vectors
     const RVector<double> x;
     const RVector<double> y;
   
     // product that I have accumulated
     double product;
   
     // constructors
     InnerProduct(const NumericVector x, const NumericVector y) 
        : x(x), y(y), product(0) {}
     InnerProduct(const InnerProduct& innerProduct, Split) 
        : x(innerProduct.x), y(innerProduct.y), product(0) {}
   
     // process just the elements of the range Ive been asked to
     void operator()(std::size_t begin, std::size_t end) {
        product += std::inner_product(x.begin() + begin, 
                                    x.begin() + end, 
                                    y.begin() + begin, 
                                    0.0);
     }
   
     // join my value with that of another InnerProduct
     void join(const InnerProduct& rhs) { 
       product += rhs.product; 
     }
   };
   
  // [[Rcpp::export]]
  double parallelInnerProduct(NumericVector x, NumericVector y) 
  {
    // declare the InnerProduct instance that takes a pointer to the vector data
    InnerProduct innerProduct(x, y);
   
    // call paralleReduce to start the work
    parallelReduce(0, x.length(), innerProduct);
   
    // return the computed product
    return innerProduct.product;
  }'
)
```

### Benchmarking

```{r ip_bench}
set.seed(1)
x <- runif(1e5)
y <- runif(1e5)

out <- microbenchmark(
  sum(x*y),
  innerProduct(x, y),
  parallelInnerProduct(x, y),
  times = 10)

autoplot(out)  
```

### Vectors of different size

```{r diff_size}
parallelInnerProduct(1:5, 1:6)
parallelInnerProduct(1:5, 1:4)
```

In serial version:

```{r diff_size2}
innerProduct(1:5, 1:4)
```

# Distance matrix calculation

In the original [post](http://gallery.rcpp.org/articles/parallel-distance-matrix/) 
the [Jensen-Shannon distance](http://en.wikipedia.org/wiki/Jensen%E2%80%93Shannon_divergence) (JSD) is used.
Here, we will simply employ the [Euclidean distance](https://en.wikipedia.org/wiki/Euclidean_distance).

## R version

```{r edistance}
edistance <- function(mat) {
  dist <- matrix(0, nrow(mat), nrow(mat))
  
  for(i in seq(1, nrow(mat) - 1)) {
    for(j in seq(i + 1, nrow(mat))) {
      sum2 <- 0 # sum of squares
      for(k in seq(1, ncol(mat))) {
        delta <- (mat[i, k] - mat[j, k])
        sum2 <- sum2 + delta * delta
      }
      
      #dist[i, j] <- sqrt(sum2)
      dist[j, i] <- sqrt(sum2) # lower triangle
    }
  }

  return(dist)
}
```

### Basic tests

All rows are equal:

```{r rows_eq}
edistance(matrix(2, 2, 2))
```

```{r rows}
edistance(matrix(1:4, 2, 2))
```

Compare with R base `dist` function (the Eucclidean distance is the default measure):

```{r comp_dist}
dist(matrix(1:9, 3, 3))
edistance(matrix(1:9, 3, 3))
```

### Clustering of iris samples

A quick scatter plot on 2 out of 4 variables:

```{r scatter_iris, fig.width = 12}
ggplot(iris, aes(Sepal.Length, Sepal.Width, color = Species)) + geom_point() +
  scale_color_brewer(type = "qual", palette = 2)
```

That gives us an idea that samples of `setosa` class are well separated from the others.

```{r clust_iris, fig.width = 12}
dat <- iris

edist <- edistance(dat[, -5])

dist <- as.dist(edist)
hc <- hclust(dist)

ord <- hc$order
lf <- data.frame(index = 1:length(ord), order = ord, class = iris[ord, 5])

p <- ggdendrogram(hc, labels = F) + 
  geom_point(data = lf, aes(x = index, y = 0, color = class)) +
  scale_colour_brewer(type = "qual", palette = 2)

p  
```

## Generic function


```{r edistance_cpp}
srcGeneric <- '
  // generic function for Euclidean distance
  template <typename InputIterator1, typename InputIterator2>
  inline double edistance(InputIterator1 begin1, InputIterator1 end1, 
    InputIterator2 begin2) 
  {
  
    // value to return
    double val = 0;
   
    // set iterators to beginning of ranges
    InputIterator1 it1 = begin1;
    InputIterator2 it2 = begin2;
   
    // for each input item
    while (it1 != end1) {
      
      // take the value and increment the iterator
      double d1 = *it1++;
      double d2 = *it2++;
      
      double delta = d2 - d1;

      val += delta * delta;
   }
   val = std::sqrt(val);

   return val;  
}'
```

## Rcpp serial version

```{r edistance_rcpp, dependson = -1}
srcSerial <- paste(
  '
  #include <Rcpp.h>
  using namespace Rcpp;
  ',
  srcGeneric,
  '
  // [[Rcpp::export]]
  NumericMatrix edistance_rcpp(NumericMatrix mat) 
  {
    // allocate the matrix we will return
    NumericMatrix rmat(mat.nrow(), mat.nrow());
   
    for(int i = 0; i < rmat.nrow(); i++) {
      for (int j = 0; j < i; j++) {
      
        // rows we will operate on
        NumericMatrix::Row row1 = mat.row(i);
        NumericMatrix::Row row2 = mat.row(j);
         
        // calculate distances
        double d = edistance(row1.begin(), row1.end(), row2.begin());
        
        // write to output matrix
        rmat(i,j) = d;
      }
    }
    return rmat;
  }
  '
)

sourceCpp(code = srcSerial)
```

## Rcpp parallel version

```{r edistance_rcpp_par}
srcParallel <- paste(
  '
  #include <Rcpp.h>
  using namespace Rcpp;
  
  // [[Rcpp::depends(RcppParallel)]]
  #include <RcppParallel.h>
  using namespace RcppParallel;

  ',
  srcGeneric,
  '
  struct EDistance : public Worker 
  {
    // input matrix to read from
    const RMatrix<double> mat;
   
    // output matrix to write to
    RMatrix<double> rmat;
   
    // initialize from Rcpp input and output matrixes (the RMatrix class
    // can be automatically converted to from the Rcpp matrix type)
    EDistance(const NumericMatrix mat, NumericMatrix rmat)
      : mat(mat), rmat(rmat) {}
    
    // function call operator that work for the specified range (begin/end)
    void operator()(std::size_t begin, std::size_t end) 
    {
      for (std::size_t i = begin; i < end; i++) {
        for (std::size_t j = 0; j < i; j++) {
          // rows we will operate on
          RMatrix<double>::Row row1 = mat.row(i);
          RMatrix<double>::Row row2 = mat.row(j);
            
          // calculate distance
          double d = edistance(row1.begin(), row1.end(), row2.begin());
          
          // write to output matrix
          rmat(i,j) = d;
        }
      }
    }
  };
  '
  ,
  '
  // [[Rcpp::export]]
  NumericMatrix edistance_rcpp_parallel(NumericMatrix mat) 
  {
    // allocate the matrix we will return
    NumericMatrix rmat(mat.nrow(), mat.nrow());

    // create the worker
    EDistance eDistance(mat, rmat);
     
    // call it with parallelFor
    parallelFor(0, mat.nrow(), eDistance);

    return rmat;
  }
  '
)

sourceCpp(code = srcParallel)
```

## Benchmarks

```{r bench_ed}
set.seed(1)
n <- 100
k <- 10
mat <- matrix(runif(n * k), ncol = k)


out <- microbenchmark(
  edistance(mat),
  dist(mat, method = "euclidean"),
  edistance_rcpp(mat),
  edistance_rcpp_parallel(mat),
  times = 10)

autoplot(out) 
```

### Excluding slow R code


```{r bench_ed2}
set.seed(1)
n <- 1000
k <- 10
mat <- matrix(runif(n * k), ncol = k)


out <- microbenchmark(
  dist(mat, method = "euclidean"),
  edistance_rcpp(mat),
  edistance_rcpp_parallel(mat),  
  times = 10)

autoplot(out) 
```

# Product of matrix and vector

## Rcpp serial version

```{r prod_mat_vec_Rcpp}
sourceCpp(code = '
  #include <Rcpp.h>
  using namespace Rcpp;

  // [[Rcpp::export]]
  NumericVector prod_mat_vec_Rcpp(NumericMatrix mat, NumericVector vec) 
  {
    // allocate the matrix we will return
    NumericVector rvec(mat.nrow());
   
    for(int i = 0; i < mat.nrow(); i++) {
      // rows we will operate on
      NumericMatrix::Row rowi = mat.row(i);
         
      // calculate inneer product
      double prod = std::inner_product(rowi.begin(), rowi.end(), vec.begin(), 0.0);
        
      // write to output matrix
      rvec(i) = prod;
    }

    return rvec;
  }
  '
)
```

## Rcpp parallel version

```{r prod_mat_vec_Rcpp_par}
sourceCpp(code = '
  #include <Rcpp.h>
  using namespace Rcpp;

  // [[Rcpp::depends(RcppParallel)]]
  #include <RcppParallel.h>
  using namespace RcppParallel;

  struct ProductMatVec : public Worker 
  {
    // input to read from
    const RMatrix<double> mat;
    const RVector<double> vec;
   
    // output to write to
    RVector<double> rvec;
   
    // initialize from Rcpp input and output matrixes (the RMatrix class
    // can be automatically converted to from the Rcpp matrix type)
    ProductMatVec(const NumericMatrix mat, const NumericVector vec, NumericVector rvec)
      : mat(mat), vec(vec), rvec(rvec) {}
    
    // function call operator that work for the specified range (begin/end)
    void operator()(std::size_t begin, std::size_t end) 
    {
      for (std::size_t i = begin; i < end; i++) {
        // rows we will operate on
        RMatrix<double>::Row rowi = mat.row(i);
        
        // calculate inneer product
        double prod = std::inner_product(rowi.begin(), rowi.end(), vec.begin(), 0.0);
        
        // write to output matrix
        rvec[i] = prod;
      }
    }
  };

  // [[Rcpp::export]]
  NumericVector prod_mat_vec_Rcpp_parallel(NumericMatrix mat, NumericVector vec) 
  {
    // allocate the matrix we will return
    NumericVector rvec(mat.nrow());
    
    // create the worker
    ProductMatVec productMatVec(mat, vec, rvec);
     
    // call it with parallelFor
    parallelFor(0, mat.nrow(), productMatVec);

    return rvec;
  }
  '
)
```


## Benchmarks

```{r bench_prod_mat_vec}
set.seed(1)
n <- 1000
k <- 100
mat <- matrix(runif(n * k), ncol = k)
vec <- runif(k) 

out <- microbenchmark(
  mat %*% vec,
  tcrossprod(mat, t(vec)),
  prod_mat_vec_Rcpp(mat, vec),
  prod_mat_vec_Rcpp_parallel(mat, vec),
  times = 20)

autoplot(out)
```

