---
title: "Rcpp: pass arguments by reference"
author: "Andrey Ziyatdinov"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: united
    toc: true
    keep_md: true
---

```{r options, echo = F}
opts_chunk$set(fig.path = "figures/", comment = NA, results = 'markup', tidy = F, message = F, warning = F, echo = T, cache = T)
```

```{r inc, echo = F, cache = FALSE}
library(plyr)
library(ggplot2)
library(gridExtra)

library(pander)
```

```{r settings, echo = F, cache = FALSE}
theme_set(theme_light())
panderOptions('table.split.table', Inf)
panderOptions('knitr.auto.asis', FALSE)
```

# About

## Problem statement

Armadillo library contains many functions that perform efficiently.
When integrating Armadillo and R via RcppArmadillo,
one can make a  mistake in how you instantiate the Armadillo object.
That in turn leads to copies and hence degraded performance.

The solution is to use an interface `const arma::vec & invec`.

See [this](http://stackoverflow.com/questions/24253228/performance-of-r-statssd-vs-armastddev-vs-rcpp-implementation) 
question for more details.

## Experiments

Here we will present both correct and missleading implementation for two operations:

* computing standard deviation in a vector of data points;
* compute matrix by vector product.

# Include

```{r include, cache = F}
library(Rcpp)

library(microbenchmark)
```

# Computing SD

Code is copied from the original [post](http://stackoverflow.com/questions/24253228/performance-of-r-statssd-vs-armastddev-vs-rcpp-implementation).


```{r Rcpp_fun}
sourceCpp(code = '
#include <RcppArmadillo.h>
// [[Rcpp::depends(RcppArmadillo)]]  

#include <vector>
#include <cmath>
#include <numeric>

// [[Rcpp::export]]
double sdCpp(Rcpp::NumericVector rinVec) 
{
  std::vector<double> inVec(rinVec.begin(), rinVec.end());
  
  int n = inVec.size();
  double sum = std::accumulate(inVec.begin(), inVec.end(), 0.0);
  double mean = sum / inVec.size();

  for(std::vector<double>::iterator iter = inVec.begin();
      iter != inVec.end(); 
      ++iter){
    double temp = (*iter - mean)*(*iter - mean);
    *iter = temp;
  }

  double sd = std::accumulate(inVec.begin(), inVec.end(), 0.0);
  return std::sqrt( sd / (n-1) );
}

// [[Rcpp::export]]      
double sdArmaByCopy(arma::colvec inVec) 
{
  return arma::stddev(inVec);
}

//  [[Rcpp::export]]    
double sdArmaByRef(const arma::colvec & inVec) 
{ 
  return arma::stddev(inVec); 
}
')
```

## Check all function computes the same

```{r sd_eq}
x <- 1:10
all(c(sdCpp(x), sdArmaByCopy(x), sdArmaByRef(x)) == sd(x))
```

## Benchmakrs

```{r sd_bench}
x <- rexp(1000)

out <- microbenchmark(sd(x), sdCpp(x), sdArmaByCopy(x), sdArmaByRef(x),
  times = 50)
  
autoplot(out)
```

## Benchmakrs

```{r sd_bench_n}
nseq <- seq(1000, 10000, length = 10)
df <- ldply(nseq, function(n) {
  x <- rexp(n)

  out <- microbenchmark(sd(x), sdCpp(x), sdArmaByCopy(x), sdArmaByRef(x),
    times = 10)
  
  df <- subset(as.data.frame(summary(out)), select = c("expr", "median"))
  df$n <- n
  
  return(df)
})  
```

```{r plot_sd_bench_n, dependson = -1}
ggplot(df, aes(n, median, color = expr)) + geom_point() + geom_line()
```

# Matrix by vector

```{r Rcpp_fun_matvec}
sourceCpp(code = '
#include <RcppArmadillo.h>
// [[Rcpp::depends(RcppArmadillo)]]  

using namespace arma;

#include <vector>
#include <cmath>
#include <numeric>

// [[Rcpp::export]]      
arma::vec prodArmaByCopy(arma::mat mat, arma::vec vec) 
{
  arma::vec rvec = mat * vec;
  return rvec;
}
 
// [[Rcpp::export]]      
arma::vec prodArmaByRef(const arma::mat & mat, const arma::vec & vec) 
{
  arma::vec rvec = mat * vec;
  return rvec;
} 
')
```

## Benchmakrs

```{r prod_bench_n}
nseq <- seq(1000, 2000, length = 10)
df <- ldply(nseq, function(n) {
  n <- ceiling(n)
  v <- rexp(n)
  M <- matrix(rexp(n*n), nrow = n, ncol = n)
  
  out <- microbenchmark(tcrossprod(M, t(v)), 
    prodArmaByCopy(M, v),
    prodArmaByRef(M, v),
    times = 10)
  
  df <- subset(as.data.frame(summary(out)), select = c("expr", "median"))
  df$n <- n
  
  return(df)
})  
```

```{r prod_sd_bench_n, dependson = -1}
ggplot(df, aes(n, median, color = expr)) + geom_point() + geom_line()
```

