---
title: "RcppEigen use cases"
author: "Andrey Ziyatdinov"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: united
    toc: true
    keep_md: true
---

```{r options, echo = F}
opts_chunk$set(fig.path = "figures/", comment = NA, results = 'markup', tidy = F, message = F, warning = F, echo = T, cache = T)
```

```{r inc, echo = F, cache = FALSE}
library(plyr)
library(ggplot2)
library(gridExtra)

library(pander)
```

```{r settings, echo = F, cache = FALSE}
theme_set(theme_light())
panderOptions('table.split.table', Inf)
panderOptions('knitr.auto.asis', FALSE)
```

# Include

```{r include, cache = F}
library(microbenchmark)

library(Rcpp)
library(RcppParallel)
```

# Column-wise storage

```{r sim_X_y}
n <- 1000
k <- 10

set.seed(1)
X <- matrix(runif(n * k), nrow = n, ncol = k)
y <- runif(k)

tX <- t(X)
ty <- t(y)
```

## Multiplication functions

### Multiplication using `%*%`

```{r mult}
str(X %*% y)
str(ty %*% tX)
```

### Multiplication using apply

```{r mult2}
mult_apply <- function(X, y) {
  prod <- rep(0, nrow(X)) 
  
  for(i in 1:ncol(X)) {
    prod <- prod + X[, i] * y[i]
  }
  
  return(prod)
}

str(mult_apply(X, y))
```

### Multiplication using `crossprod`

```{r mult_crossprod}
str(tcrossprod(X, ty))
str(tcrossprod(ty, X))
```

```{r mult_crossprod2}
str(crossprod(tX, y))
```

### Multiplication using `mult_RcppEigen`

```{r mult_RcppEigen, cache = F}
sourceCpp("src/mult_RcppEigen.cpp")
```

```{r mult_RcppEigen2}
str(mult_RcppEigen(X, y))
```

### Multiplication using `mult_apply_RcppEigen_colwise`

```{r mult_apply_RcppEigen_colwise, cache = F}
sourceCpp("src/mult_apply_RcppEigen_colwise.cpp") 
```

```{r mult_apply_RcppEigen_colwise2}
str(mult_apply_RcppEigen_colwise(X, y))
```

### Multiplication using `mult_apply_RcppEigen_rowwise`

```{r mult_apply_RcppEigen_rowwise, cache = F}
sourceCpp("src/mult_apply_RcppEigen_rowwise.cpp") 
```

```{r mult_apply_RcppEigen_rowwise2}
str(mult_apply_RcppEigen_rowwise(X, y))
```

### Multiplication using `mult_apply_Rcpp_rowwise`

```{r mult_apply_Rcpp_rowwise, cache = F}
sourceCpp("src/mult_apply_Rcpp_rowwise.cpp") 
```

```{r mult_apply_Rcpp_rowwise2}
str(mult_apply_Rcpp_rowwise(X, y))
```

### Multiplication using `par_mult_apply_Rcpp_rowwise`

```{r par_mult_apply_Rcpp_colwise, cache = F}
sourceCpp("src/par_mult_apply_Rcpp_colwise.cpp") 
```

```{r par_mult_apply_Rcpp_colwise2}
str(par_mult_apply_Rcpp_colwise(X, y, chunkSize = 1, verbose = 1))

str(par_mult_apply_Rcpp_colwise(X, y, chunkSize = ncol(X), verbose = 1))
```


#### Wrapper `par_mult_apply_Rcpp_rowwise`


```{r wrap_par_mult_apply_Rcpp_colwise}
wrap_par_mult_apply_Rcpp_colwise <- function(..., cores = -1, chunkSize = 1)
{
  stopifnot(require(RcppParallel))
  
  if(!missing(cores)) { 
    RcppParallel::setThreadOptions(numThreads = cores)
  }
  
  par_mult_apply_Rcpp_colwise(..., chunkSize = chunkSize)
}
```


### Multiplication using `mult_apply_Rcpp_rowwise`

```{r mult_apply_Rcpp_colwise, cache = F}
sourceCpp("src/mult_apply_Rcpp_colwise.cpp") 
```

```{r mult_apply_Rcpp_colwise2}
str(mult_apply_Rcpp_colwise(X, y))
```


## Benchmarks (k << n)

```{r bench}
nseq <- seq(100, 2000, length = 5)
df <- ldply(nseq, function(n) {
  n <- ceiling(n)
  k <- 100

  set.seed(1)
  X <- matrix(runif(n * k), nrow = n, ncol = k)
  y <- runif(k)

  tX <- t(X)
  ty <- t(y)
  
  cat(" * n:", n, "\n")
  
  out <- microbenchmark(
    X %*% y,
    mult_apply(X, y),
    mult_apply_RcppEigen_colwise(X, y),
    mult_apply_RcppEigen_rowwise(X, y),
    mult_apply_Rcpp_rowwise(X, y), 
    mult_apply_Rcpp_colwise(X, y, vectorize = TRUE),
    mult_apply_Rcpp_colwise(X, y),       
    mult_RcppEigen(X, y),
    tcrossprod(X, ty), 
    crossprod(tX, y),
    times = 10)
  
  df <- subset(as.data.frame(summary(out)), select = c("expr", "median"))
  df$n <- n
  
  return(df)
})  
```

```{r bench_plot, dependson = -1}
ggplot(df, aes(n, median, color = expr)) + geom_point() + geom_line()
``` 

```{r bench_plot_subset, dependson = -2}
ggplot(subset(df, n > 1000 & (expr %in% c("mult_apply_RcppEigen_colwise(X, y)",
  "mult_apply_RcppEigen_rowwise(X, y)", "mult_apply_Rcpp_rowwise(X, y)",
  "mult_apply_Rcpp_colwise(X, y, vectorize = TRUE)",
  "mult_apply_Rcpp_colwise(X, y)", "mult_RcppEigen(X, y)"))), 
  aes(n, median, color = expr)) + geom_point() + geom_line()
```

```{r bench_plot_subset2, dependson = -3}
ggplot(subset(df, n > 1000 & (expr %in% 
  c("mult_apply_RcppEigen_colwise(X, y)", "mult_apply_Rcpp_colwise(X, y)",
    "mult_apply_RcppEigen_rowwise(X, y)",
    "mult_RcppEigen(X, y)"))), 
  aes(n, median, color = expr)) + geom_point() + geom_line()
```


## Benchmarks (k == n)

```{r bench2}
nseq <- seq(500, 5000, length = 5)
df <- ldply(nseq, function(n) {
  n <- ceiling(n)
  k <- n

  set.seed(1)
  X <- matrix(runif(n * k), nrow = n, ncol = k)
  y <- runif(k)

  tX <- t(X)
  ty <- t(y)
  
  cat(" * n:", n, "\n")
  
  out <- microbenchmark(
    X %*% y,
    mult_apply_RcppEigen_colwise(X, y),
    mult_apply_RcppEigen_rowwise(X, y),
    mult_RcppEigen(X, y),
    tcrossprod(X, ty), 
    times = 10)
  
  df <- subset(as.data.frame(summary(out)), select = c("expr", "median"))
  df$n <- n
  
  return(df)
})  
```

```{r bench2_plot, dependson = -1}
ggplot(df, aes(n, median, color = expr)) + geom_point() + geom_line()
```

```{r bench2_plot_2, dependson = -2}
ggplot(subset(df, n > 1000), aes(n, median, color = expr)) + geom_point() + geom_line()
```

```{r bench2_plot_3, dependson = -3}
ggplot(subset(df, n > 1000 & !(expr %in% c("X %*% y", "mult_apply_RcppEigen_rowwise(X, y)"))), 
  aes(n, median, color = expr)) + geom_point() + geom_line()
```

## Benchmarks (k == n, parallel)

```{r bench3}
nseq <- seq(500, 5000, length = 5)
df <- ldply(nseq, function(n) {
  n <- ceiling(n)
  k <- n

  set.seed(1)
  X <- matrix(runif(n * k), nrow = n, ncol = k)
  y <- runif(k)

  tX <- t(X)
  ty <- t(y)
  
  cat(" * n:", n, "\n")
  
  out <- microbenchmark(
    mult_RcppEigen(X, y),
    mult_apply_Rcpp_colwise(X, y), 
    wrap_par_mult_apply_Rcpp_colwise(X, y, chunkSize = n, cores = 1),
    wrap_par_mult_apply_Rcpp_colwise(X, y, cores = 2),  
    times = 10)
  
  df <- subset(as.data.frame(summary(out)), select = c("expr", "median"))
  df$n <- n
  
  return(df)
})  
```


```{r bench3_plot, dependson = -1}
ggplot(df, aes(n, median, color = expr)) + geom_point() + geom_line()
```

```{r bench3_plot_2, dependson = -2}
ggplot(subset(df, n > 1000), aes(n, median, color = expr)) + geom_point() + geom_line()
```

### Benchmarks (k == n, parallel, larger matrices)

```{r bench4}
nseq <- seq(3000, 7000, length = 5)
df <- ldply(nseq, function(n) {
  n <- ceiling(n)
  k <- n

  set.seed(1)
  X <- matrix(runif(n * k), nrow = n, ncol = k)
  y <- runif(k)

  tX <- t(X)
  ty <- t(y)
  
  cat(" * n:", n, "\n")
  
  out <- microbenchmark(
    mult_RcppEigen(X, y),
    wrap_par_mult_apply_Rcpp_colwise(X, y, chunkSize = 100, cores = 2),  
    times = 10)
  
  df <- subset(as.data.frame(summary(out)), select = c("expr", "median"))
  df$n <- n
  
  return(df)
})  
```


```{r bench4_plot, dependson = -1}
ggplot(df, aes(n, median, color = expr)) + geom_point() + geom_line()
```

